---
title: "Introduction to FastBreakR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to FastBreakR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Welcome to the FastBreakR package! FastBreakR is a fast, efficient, and clean way to fetch data from the NBA API. Designed for analysts, modelers, and basketball enthusiasts, FastBreakR simplifies the process of acquiring, consolidating, and analyzing NBA statistics. While the package is already functional with its core set of tools, it remains under development, and new features are continually being added.

This vignette introduces some the primary functions of FastBreakR, demonstrating their capabilities and providing examples to help you get started.

## Key Features

- **Easy-to-use functions** for fetching and processing various types of NBA data,
including game scores, team statistics, shot data and play by play data.
- **Modular design** that seamlessly integrates NBA data into your analysis pipeline.
- **Integration with `tidyverse`**, enabling effortless manipulation, analysis, and visualization of NBA data.

## FastBreakR: Built for Efficiency and Flexibility

FastBreakR is designed to be both efficient and flexible. It’s user-friendly for beginners while offering advanced capabilities for experienced users who want to dive deeper into NBA statistics. Whether you're building a sports analytics app, conducting research, or simply exploring basketball data, FastBreakR is your go-to solution.

## Prerequisites
Before you start, ensure you have installed the FastBreakR package and its dependencies. You can install FastBreakR directly from GitHub:

```r
# Install from GitHub
devtools::install_github("Jay-Kay-15/FastBreakR")
install.packages("dplyr")
```

## Getting Started

To get started with FastBreakR, simply load the package:

```{r, load packages}
# Load the FastBreakR package
library(FastBreakR)
library(dplyr, warn.conflicts = FALSE)

```

## Fetching Team Stats

The **`nba_team_stats()`** function retrieves NBA team statistics for the seasons you specify. The output is a list, where each element corresponds to a season and contains five data frames with various stat types, including:

- Traditional Stats
- Advanced Stats
- Four Factors
- Miscellaneous Stats
- Scoring Stats


```{r, team stats}

# Fetch team stats for the 2022-2023 & 2023-2024 regular season
team_stats <- nba_team_stats(2023:2024, season_type = "Regular Season")

# Display a glimpse of the advanced stats for the 2024 season
glimpse(team_stats$season_2024$Advanced)


```

Once you have the data, you can analyze it as is or consolidate it using additional package functions.

## Additional Functions for Cleaning and Processing Data

Once you've fetched the data, you might want to clean and process it further to meet your analysis requirements. MambaNBA provides functions such as **`consolidate_stats()`** and **`add_schedule_details()`** to help you with these tasks.

For example, you can consolidate team stats and calculate additional metrics based on the schedule:

```{r, process team stats}

# Process team stats using the data from scrape_nba_team_stats()
team_stats_processed <- team_stats %>%
    consolidate_stats() %>%      # Consolidates stats into one data frame
    add_schedule_details()      # Calculates additional schedule metrics

# Display a glimpse of the processed data
glimpse(team_stats_processed$season_2024)

```

**`consolidate_stats()`**: This function combines data from multiple statistical categories (e.g., basic, advanced, scoring) into a single data frame for easier analysis.

**`add_schedule_details()`**: This function calculates metrics related to a team's schedule, such as back-to-back games, rest days, and more.

These functions are particularly useful for cleaning, transforming, and preparing the data for deeper insights or more advanced modeling.

## Fetching Box Scores

The **`nba_box_scores()`** function allows you to fetch detailed box score data for specified games and stat categories. The output is a data frame with all duplicate columns between
stat categories removed.

```{r, fetch box scores, error = FALSE}

# Here, we're fetching play-by-play data for a subset of game IDs
games <- team_stats_processed$season_2024 %>%
    filter(game_date == "2023-12-25") %>%
    pull(game_id) %>%
    unique()

# Fetching Traditional and Miscellaneous box scores
box_scores <- nba_box_scores(
  measure_types = c("Traditional", "Misc"),
  game_ids = games
)

# Display a glimpse of the processed data
glimpse(box_scores)

```

## Fetching Play-by-Play Data with Rate Limiting

Fetching play-by-play data from the NBA can be resource-intensive, especially when requesting data for many games in quick succession. To ensure that your requests don't result in timeouts or errors due to rate limiting, FastBreakR incorporates an automatic rate-limiting mechanism.

Below is an example of how to fetch play-by-play data for a specific set of games, using the **`nba_play_by_play()`** function. This example demonstrates how to fetch data for a list of game IDs, while respecting rate limits to avoid overloading the server.

```{r, fetch play by play}

# Fetch the play-by-play data for the same games
play_by_play_data <- nba_play_by_play(games)

# Display a glimpse of the fetched data
glimpse(play_by_play_data)

```

In this example:

- We fetch play-by-play data for a small subset of game IDs (from game_list).
- The **`scrape_nba_play_by_play()`** function handles the rate-limiting process automatically, ensuring the requests are spaced out to avoid server timeouts.
- The returned data is a dataframe containing the play-by-play events for the games, which you can analyze or visualize further.

## Conclusion

FastBreakR is your go-to tool for NBA data analysis, offering a robust and growing set of features. Whether you are creating models, performing analyses, or exploring trends, this package streamlines the process, enabling you to focus on insights instead of data fetching.

Stay tuned for updates as we continue to expand and improve FastBreakR. Happy analyzing!







